---
title: "Data for Stat 433"
author: "Celina Ough, Maya Tunney, Robert George"
output: html_document
---
### Library funding and demographics by Wisconsin county

This Census Data contains information on the breakdown of household types and income for counties in Wisconsin. For example, this data includes the median income for families, single people, and married couples in the county, as well as median income by race. 
```{r, echo = FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(stringr)
```

```{r}
income_2019 <- read_csv("productDownload_2021-10-25T203440/ACSST1Y2019.S1903_data_with_overlays_2021-10-25T203421.csv", 
    na = "N",skip=1)%>%
  select(-id)%>%
  select(-matches("Margin"),-matches("Distribution"),-matches("FAMILY SIZE"))
colnames(income_2019) = gsub('Estimate!!Number!!HOUSEHOLD INCOME BY RACE AND HISPANIC OR LATINO ORIGIN OF HOUSEHOLDER!!', '', colnames(income_2019), fixed=TRUE)
colnames(income_2019) = gsub('Estimate!!Number!!HOUSEHOLD INCOME BY AGE OF HOUSEHOLDER!!', '', colnames(income_2019), fixed=TRUE)
colnames(income_2019) = gsub('Estimate!!Number!!FAMILIES!!', '', colnames(income_2019), fixed=TRUE)
colnames(income_2019) = gsub('Estimate!!Number!!NONFAMILY HOUSEHOLDS!!', '', colnames(income_2019), fixed=TRUE)
colnames(income_2019) = gsub('Estimate!!Number!!', '', colnames(income_2019), fixed=TRUE)
head(income_2019)
```

This dataset includes information about Public Libraries in Wisconsin. It includes information about funding, location, hours, resources and how the public utilizes the library. 
```{r}
library(readxl)
libraries_2019 <- read_excel("PRELIMINARY_2019_public_library_service_data.xlsx", na = "-", skip = 1)[-1,]
head(libraries_2019)
```
