---
title: "Main Project"
author: "Celina"
date: "10/28/2021"
output: html_document
---

```{r, include=FALSE}
# Necessary Packages 
rm(list=ls())
library(readr)
library(tidyverse)
library(dplyr)
library(stringr)
library(maps)
library(ggplot2)
library(boot)
library(car)
data(countyMapEnv)
```


# Downloading the Data and joining  
```{r, message=FALSE, warning=FALSE}
#Demographic Data
demo_2019 <- read_csv("data\\WICounty2019Demographics.csv", skip = 1) %>%
  select(-matches("(Margin)|(Percent)|(ratio)|(_1)"), -id) %>%
  select(matches("(Total population)|(Geographic Area)")) %>%
  select(-matches("Hispanic or Latino \\(of any race\\)!!"), 
         -matches("American Indian and Alaska Native!!"),
         -matches("Asian!!"),
         -matches("Native Hawaiian and Other Pacific Islander!!"),
         -matches("Two or more races!"), 
         -matches("and over$"), 
         -matches("Race alone or in combination with one or more other races!!")) %>% 
  rename_all(funs(str_replace_all(., "((Estimate!!)|(Total population!!)+)", ""))) %>%
  rename(County = "Geographic Area Name") %>%
  mutate(County = sub(" County, Wisconsin", "", County))

#Income Data
income_2019 <- read_csv("data\\WICounty2019Income.csv", na = "N", skip=1) %>%
  select(-id,-matches("(Margin)|(Distribution)")) %>%
  select(-matches("race"), -matches("FAMILIES!!families!!")) %>%
  rename_all(funs(str_replace_all(., "((Estimate!!)|(Number!!))+", ""))) %>%
  rename(County = "Geographic Area Name")%>%
  mutate(County = sub(" County, Wisconsin", "", County))

#Library Data
library_data <- readxl::read_excel("data\\public_library_service_data_2019.xlsx", sheet = "2019 PRELIMINARY DATA", skip = 1) %>%
  slice(-1) %>%
  select(-c(
            "Library ID", 
            "Municipality",
            "Public Library System",
            "Branches", 
            "15. No. of Bookmobiles Owned", 
            "Other Service Outlets", 
            "Books-by-Mail", 
            "Joint Library under s.43.53 1=Yes", 
            "Participating Municipalities", 
            "Circulation Count Method: Actual / Survey", 
            "Locale Code", 
            "Other Material Description"), 
         -matches("(Outlay)|(Access Denied)|(Databases)|(Circulation)|(Collection Retrievals)")) 

data <- library_data %>%
  na_if("0") %>%
  left_join(demo_2019, by="County") %>%
  inner_join(income_2019, by="County") 


 programs_staff <- data %>%
  select(County, matches("(Number .*Program)|(Staff)"), "Total Income") %>%
  drop_na() %>%
  rename_all(funs(str_replace_all(., " ", "_"))) %>%
  mutate(across(.cols=-c(County), as.numeric)) %>%
  group_by(County) %>%
  mutate(across(.cols=everything(), sum)) %>%
  unique() 


library_income_costs <- data %>%
  select(County, "Total Income", 
         "Salaries & Wages", 
         "Print Materials", 
         "Electronic format", 
         "Audiovisual Materials", 
         "Contracted Services", 
         "Other Operating Expenditures", 
         "Total Operating Expenditures", 
         "Resident Support per Capita (Local Revenues and Resident Population)") %>%
  rename(Library_Income = `Total Income`) %>%
  rename_all(funs(str_replace_all(., " ", "_"))) %>%
  rename(Resident_Support_per_Capita = `Resident_Support_per_Capita_(Local_Revenues_and_Resident_Population)`) %>%
  drop_na() %>%
  mutate(across(.cols=-c(County), as.numeric)) %>%
  group_by(County) %>%
  mutate(across(.cols=everything(), sum)) %>%
  unique()

# Note: Library focuses activies on Children (0-11), Young Adults (12-18), and other. 
# They only recorded the Number of Uses of computers, they didn't divide it based on if they had access to Wifi
# Not all locations record Wireless Uses or Library Visits but, I thought they could be useful. 

computer_Usage_Data <- data %>%
  select(County, 
         matches("SEX AND AGE"), 
         -matches("(Male)|(Female)|(Median)"), 
         "Median income (dollars)!!FAMILIES!!Families", 
         "Median income (dollars)!!NONFAMILY HOUSEHOLDS!!Nonfamily households", 
          matches("Computers"), "Wireless Internet Uses", 
         "Library Visits") %>%
  drop_na() %>%
  rename(Total_Population = "SEX AND AGE!!Total population", 
         Income_Families = "Median income (dollars)!!FAMILIES!!Families", 
         Income_NonFamily = "Median income (dollars)!!NONFAMILY HOUSEHOLDS!!Nonfamily households") %>%
  rename_all(funs(str_replace_all(., "((SEX AND AGE!!)|( years))+", ""))) %>%
  rename_all(funs(str_replace_all(., " ", "_"))) %>%
  mutate(across(.cols=-c(County), as.numeric)) %>%
  group_by(County)  %>%
  mutate(across(.cols=c("Number_of_Public_Use_Computers", "Number_of_Public_Use_Computers_with_Internet_Access", "Uses_of_Public_Internet_Computers", "Wireless_Internet_Uses", "Library_Visits"), sum), 
         Uses_per_Computer = Uses_of_Public_Internet_Computers/Number_of_Public_Use_Computers, 
         Uses_per_Computer_with_Internet = Uses_of_Public_Internet_Computers/Number_of_Public_Use_Computers_with_Internet_Access)%>%
  unique() 

```




#### General Introduction: Understanding the data and where funding goes

________________________ **READ ME** ______________________________________

I've created three new datasets: 
1. `programs_staff` = Number of Programs and Staff per County 
2. `Library_income_costs` = Total Income and Costs for library (All Numbers here are in Dollars)  
3. `computer_usage_data` = All Computer Usage by county (column names explain more)

Notes for all three datasets: 
1. I've renamed all the columns to remove spaces so we don't have to use ` anymore and RStudio can autocomplete the variables  
2. I've dropped all the rows that contain missind data for each new dataset. We can add that back in later if we want but, for now they are dropped. 
3. All numbers in the new datasets are actual numbers and not character strings so you can use math functions on them. 



# Basic Graphs of Computer Usage with Income Levels 


```{r}
computer_usage_graph <- computer_Usage_Data %>%
  left_join(library_income_costs, by="County") %>%
  select("Total_Population", "Uses_per_Computer", "Library_Visits", "Income_Families", "Wireless_Internet_Uses", "Income_NonFamily", "County") %>%
  pivot_longer(cols = c("Income_Families", "Income_NonFamily"), names_to = "Income_Type", values_to = "Income")

computer_usage_graph %>%
  ggplot(aes(x = Income, y = Uses_per_Computer/Total_Population)) + 
  geom_point() + 
  labs(y = "Ratio of Library Computer Usage to County Population", x = "Income", title = "Relationship of County Wealth to Library Computer Usage") + 
  #geom_smooth(method = lm, se=FALSE, color="red") + 
  facet_wrap(~Income_Type)

computer_usage_graph %>%
  ggplot(aes(x = Income, y = Library_Visits/Total_Population)) + 
  geom_point() + 
  labs(x = "Income", y = "Number of Library Visits", title = "Library Visits v. County Population") + 
  #geom_smooth(method = lm, se=FALSE, color="red") + 
  facet_wrap(~Income_Type)

computer_usage_graph %>%
  ggplot(aes(x = Income, y = Wireless_Internet_Uses/Total_Population)) + 
  geom_point() + 
  labs(x = "Income", y = "Wireless Usage", title = "Wireless Usage v. County Population") + 
  #geom_smooth(method = lm, se=FALSE, color="red") + 
  facet_wrap(~Income_Type)
```

### Outliers  
```{r}
# Some outlining data points 

# Dane County Median Income is MUCH HIGHER than other places 
computer_Usage_Data %>% 
  select(matches("ncome")) %>%
  filter(Income_Families > 1500000)

computer_usage_graph %>%
  filter(Wireless_Internet_Uses/Total_Population > 1)
```


### Where Library Income Goes 
```{r}
library_income_costs %>%
  select(-c(Library_Income, Total_Operating_Expenditures, Resident_Support_per_Capita)) %>%
  mutate(Materials = sum(Print_Materials, Electronic_format, Audiovisual_Materials)) %>%
  select(`Salaries_&_Wages`, Materials, Contracted_Services, Other_Operating_Expenditures, County) %>%
  mutate(Total = sum(`Salaries_&_Wages`, Materials, Contracted_Services, Other_Operating_Expenditures)) %>%
  summarise(`Salaries & Wages` = `Salaries_&_Wages`/Total, 
            Materials = Materials/Total, 
            Contracted_Services = Contracted_Services/Total, 
            Other = Other_Operating_Expenditures/Total) %>%
  
  pivot_longer(cols = matches("a")) %>%
  ggplot(aes(x = name, y = value)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+ 
  facet_wrap(~County)

# Most income primarily goes towards paying salaries. What's the average salary per person? 

programs_staff %>% 
  left_join(library_income_costs, by = "County") %>%
  select(County, Total_Staff, `Salaries_&_Wages`, Other_Paid_Staff) %>%
  drop_na() %>%
  summarise(Ave_Salary = `Salaries_&_Wages`/Total_Staff, 
            non_librarian = Other_Paid_Staff/Total_Staff) %>%
  arrange(Ave_Salary) %>%
  left_join(computer_Usage_Data, by = "County") %>%
  select(County, Ave_Salary, Income_Families, non_librarian) %>%
  ggplot(aes(x = Income_Families, y = Ave_Salary)) + 
  geom_point() + 
  labs(x = "Median Family Income", y = "Average Staff Salary", title = "Average Salary by County Wealth")

```



### Income Maps 
```{r}
plot_map <- computer_Usage_Data %>%
  mutate(County = tolower(County))

map_data("county")%>%
  filter(region == "wisconsin")%>%
  left_join(plot_map, by=c("subregion"="County")) %>% 
  ggplot(aes(x = long, y = lat, outline = subregion, fill = Income_NonFamily)) +
  geom_polygon(colour = "grey50") +
  scale_fill_continuous(type = "viridis", na.value = "grey80") + 
  theme(axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank()) + 
  labs(title = "Median Non-Family Income for Wisconsin Counties", fill = "Median Non-Family \nIncome (dollars)")

map_data("county")%>%
  filter(region == "wisconsin")%>%
  left_join(plot_map, by=c("subregion"="County")) %>% 
  ggplot(aes(x = long, y = lat, outline = subregion, fill = Income_Families)) +
  geom_polygon(colour = "grey50") +
  scale_fill_continuous(type = "viridis", na.value = "grey80") + 
  theme(axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank()) + 
  labs(title = "Median Family Income for Wisconsin Counties", fill = "Median Family \nIncome (dollars)")

```

```{r}
data2 <- data %>%
  select(County, 
         `SEX AND AGE!!Total population`, 
         `SEX AND AGE!!Median age (years)`, 
         `Median income (dollars)!!FAMILIES!!Families`, 
         `Median income (dollars)!!NONFAMILY HOUSEHOLDS!!Nonfamily households`, 
         `Library Visits`, 
         `Total Income`, 
         `Number of Public Use Computers`, 
         `Uses of Public Internet Computers`) %>%
  drop_na() %>%
  mutate(across(.cols=-c(County), as.numeric), 
         County = tolower(County)) %>%
  rename(Total_Population = (`SEX AND AGE!!Total population`), 
         Median_Age = (`SEX AND AGE!!Median age (years)`), 
         Family_Income = `Median income (dollars)!!FAMILIES!!Families`, 
         NonFamily_Income = `Median income (dollars)!!NONFAMILY HOUSEHOLDS!!Nonfamily households`) %>%
  group_by(County) %>%
  mutate(num_libraries = n(), 
         Library_Visits = sum(`Library Visits`), 
         Library_Income = sum(`Total Income`), 
         Num_Computers = sum(`Number of Public Use Computers`), 
         Computer_Usage = sum(`Uses of Public Internet Computers`)) %>%
  select(-matches(" ")) %>%
  unique()


ggplot(data2, aes(Family_Income/1000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Family Income (1k dollars)", y = "Count", title = "Median Family Income per County") +
  scale_x_continuous(breaks = seq(0, 120, 5))

ggplot(data2, aes(NonFamily_Income/1000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Non-Family Income (1k dollars)", y = "Count", title = "Median Non-Family Income per County") +
  scale_x_continuous(breaks = seq(0, 60, 2.5))

ggplot(data2, aes(Library_Income/num_libraries/100000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Library Funding (100k dollars)", y = "Count", title = "Average Library Funding for each County") 

ggplot(data2, aes(Total_Population/num_libraries/1000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Population per Library (1k)", y = "Count", title = "Average Population Served by Libary for each County") + 
  scale_x_continuous(breaks = seq(0, 150, 25))

ggplot(data2, aes(Computer_Usage/Num_Computers/Total_Population)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Computer Usage", y = "Count", title = "Computer Usages per County", subtitle = "Normalized by Number of Computers and County Population")# +
  scale_x_continuous(breaks = seq(0, 600, 100))

ggplot(data2, aes(Total_Population/Num_Computers)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Population Served by each Computer", y = "Count", title = "Average Population Served by each Computer") +
  scale_x_continuous(breaks = seq(0, 3000, 500))

ggplot(data2, aes(Median_Age)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Median County Age", y = "Count", title = "Median Age by County") +
  scale_x_continuous(breaks = seq(0, 46, 1))

ggplot(data2, aes(Total_Population/1000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "County Population (1k Residents)", y = "Count", title = "County Populations") +
  scale_x_continuous(breaks = seq(0, 1000, 100))

ggplot(data2, aes(Library_Visits/Total_Population)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Library Visits/County Population", y = "Count", title = "Library Visits") 

ggplot(data2, aes(Library_Income/Total_Population)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Library Income/County Population", y = "Count", title = "Library Income as a Proportion of Population")
ggplot(data2, aes(x = Library_Income, y = Total_Population)) + 
  geom_point() 

ggplot(data2) + 
  geom_point(aes(x = Library_Income, y = NonFamily_Income, colour = "Non-Family Income")) + 
  geom_point(aes(x = Library_Income, y = Family_Income, colour = "Family Inncome")) + 
  labs(x = "Library Funding", y = "Median Income", title = "The Relationship between Income and Library Funding")
```

```{r}

# RELATIONSHIPS

ggplot(data2, aes(Total_Population/Num_Computers)) +
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Population Served by each Computer", y = "Count", title = "Average Population Served by each Computer") +
  scale_x_continuous(breaks = seq(0, 3000, 500))

ggplot(data2, aes(y = Total_Population, x = Num_Computers)) + 
  geom_point() 

ggplot(data2, aes(Computer_Usage/Num_Computers/Total_Population)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Computer Usage", y = "Count", title = "Computer Usages per County", subtitle = "Normalized by Number of Computers and County Population")# +
  scale_x_continuous(breaks = seq(0, 600, 100))

# COUNTY DATA EXPLORATION

ggplot(data2, aes(Total_Population/1000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "County Population (1k Residents)", y = "Count", title = "County Populations") +
  scale_x_continuous(breaks = seq(0, 1000, 100))

map_data("county")%>%
  filter(region == "wisconsin")%>%
  left_join(data2, by=c("subregion"="County")) %>% 
  ggplot(aes(x = long, y = lat, outline = subregion, fill = Total_Population)) +
  geom_polygon(colour = "grey50") +
  scale_fill_continuous(type = "viridis", na.value = "grey80") + 
  theme(axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank()) + 
  labs(title = "Total Population for Wisconsin Counties", fill = "Total Population")

map_data("county")%>%
  filter(region == "wisconsin")%>%
  left_join(data2, by=c("subregion"="County")) %>% 
  ggplot(aes(x = long, y = lat, outline = subregion, fill = Median_Age)) +
  geom_polygon(colour = "grey50") +
  scale_fill_continuous(type = "viridis", na.value = "grey80") + 
  theme(axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank()) + 
  labs(title = "Median Age for Wisconsin Counties", fill = "Median Age")

map_data("county")%>%
  filter(region == "wisconsin")%>%
  left_join(data2, by=c("subregion"="County")) %>% 
  ggplot(aes(x = long, y = lat, outline = subregion, fill = Family_Income)) +
  geom_polygon(colour = "grey50") +
  scale_fill_continuous(type = "viridis", na.value = "grey80") + 
  theme(axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank()) + 
  labs(title = "Median Family Income for Wisconsin Counties", fill = "Median Family \nIncome (dollars)")


# LIBRARY DATA EXPLORATION

ggplot(data2, aes(x = Total_Population, y = Library_Visits)) + 
  geom_point() + 
  labs(x = "Library Visits", y = "County Population", title = "Library Visits vs. County Population")

ggplot(data2, aes(y = Library_Income, x = Total_Population)) + 
  geom_point() + 
  labs(x = "Library Income", y = "County Population", title = "Library Income vs. County Population")

ggplot(data2) + 
  geom_point(aes(x = Library_Income, y = NonFamily_Income, colour = "Non-Family Income")) + 
  geom_point(aes(x = Library_Income, y = Family_Income, colour = "Family Inncome")) + 
  labs(x = "Library Funding", y = "Median Income", title = "The Relationship between Income and Library Funding")

ggplot(data2, aes(Library_Income/num_libraries/100000)) + 
  geom_histogram(color = "black", fill = "grey") + 
  labs(x = "Library Funding (100k dollars)", y = "Count", title = "Average Library Funding for each County") 
```



```{r}
# runExample("01_hello")      # a histogram
# runExample("02_text")       # tables and data frames
# runExample("03_reactivity") # a reactive expression
runExample("04_mpg")        # global variables
# runExample("05_sliders")    # slider bars
runExample("06_tabsets")    # tabbed panels
# runExample("07_widgets")    # help text and submit buttons
```

```{r}
#MODELING RELATIONSHIP BETWEEN COMPUTER USAGE AND INCOME
meanusage = mean(computer_Usage_Data$Uses_per_Computer_with_Internet)
ogfit = glm(Uses_per_Computer_with_Internet<meanusage~Income_Families+Income_NonFamily, data = computer_Usage_Data)
summary(ogfit)
#It seems to explain a good amount of deviance as well.
Anova(glm(Uses_per_Computer_with_Internet<meanusage~Income_Families+Income_NonFamily, data = computer_Usage_Data))

#Train/Test split
dt = sort(sample(nrow(computer_Usage_Data), nrow(computer_Usage_Data)*.7))
train<-computer_Usage_Data[dt,]
test<-computer_Usage_Data[-dt,]
fit = glm(Uses_per_Computer_with_Internet<meanusage~Income_Families+Income_NonFamily, data = train)
pred = predict(fit, test, type="response")>0.5
actual = test$Uses_per_Computer_with_Internet<meanusage
mean(pred == actual)

#Cross-Validation: raw and adjusted CV estimates of prediction error using 6-fold validation
cv.glm(computer_Usage_Data, ogfit, K = 6)$delta

#Let's look at a plot:
plot_axes = expand.grid(Income_Families = seq(from = 67929, to = 113591, by = (113591-67929)/100),
                        Income_NonFamily = seq(from = 31486, to = 51106, by = (51106-31486)/100))
plot_axes$pred = plogis(predict(ogfit, newdata = plot_axes))
ggplot(plot_axes)+
  geom_tile(aes(x = Income_Families, y = Income_NonFamily, fill = pred))+
  scale_fill_gradient(low = "white", high = "blue") + 
  labs(x = "Family Income", y = "Non-Family Income", title = "Predicted Effects of Family vs Non-Family Income on Library Computer Usage")
```
We can see that our model has about 25% prediction error when we do 6 different random train/test samples. Also, interestingly, our plot shows that at all levels of family income, as nonfamily income diminishes, usage per computer is more likely to be above average. The opposite effect is true for family income.


