---
title: "DataSource"
author: "Celina Ough, Maya Tunney, Robert George"
date: "10/25/2021"
output: html_document
---

## Short description of County Data  

This is census data with Demographic data from counties in Wisconsin. It includes information on race, age, and sex. 

This Census Data contains information on the breakdown of household types and income for counties in Wisconsin. For example, this data includes the median income for families, single people, and married couples in the county. 

This dataset includes information about Public Libraries in Wisconsin. It includes information about funding, location, hours, resources and how the public utilizes the library. 

## Short description of Library Data 

TBD

```{r, include=FALSE}
rm(list=ls())
library(readr)
library(tidyverse)
library(dplyr)
library(stringr)
```


```{r, message=FALSE, warning=FALSE}
WI_data <- read_csv("..\\data\\WICounty2019Demographics.csv", skip = 1) %>%
  select(-matches("Margin"), 
         -matches("Percent"), 
         -matches("ratio"), 
         -id)

head(WI_data)
```


```{r}
library_data <- readxl::read_excel("..\\data\\public_library_service_data_2019.xlsx", sheet = "2019 PRELIMINARY DATA", skip = 1) %>%
  slice(-1) %>%
  select(-c("Library ID", "Branches", "15. No. of Bookmobiles Owned", 
            "Other Service Outlets", "Books-by-Mail", "Joint Library under s.43.53 1=Yes", 
            "Circulation Count Method: Actual / Survey", "Locale Code"), 
         -matches("Outlay"), -matches("Access Denied"), -matches("Databases"))

head(library_data)
```



## Robert's Code 

### Library funding and demographics by Wisconsin county

This Census Data contains information on the breakdown of household types and income for counties in Wisconsin. For example, this data includes the median income for families, single people, and married couples in the county, as well as median income by race. 
```{r, echo = FALSE, include=FALSE}
# library(readr)
# library(tidyverse)
# library(dplyr)
# library(stringr)
```

```{r, include=FALSE}
# income_2019 <- read_csv("..\\data\\WICounty2019Demographics.csv", 
#     na = "N",skip=1)%>%
#   select(-id)%>%
#   select(-matches("Margin"),-matches("Distribution"),-matches("FAMILY SIZE"))
# colnames(income_2019) = gsub('Estimate!!Number!!HOUSEHOLD INCOME BY RACE AND HISPANIC OR LATINO ORIGIN OF HOUSEHOLDER!!', '', colnames(income_2019), fixed=TRUE)
# colnames(income_2019) = gsub('Estimate!!Number!!HOUSEHOLD INCOME BY AGE OF HOUSEHOLDER!!', '', colnames(income_2019), fixed=TRUE)
# colnames(income_2019) = gsub('Estimate!!Number!!FAMILIES!!', '', colnames(income_2019), fixed=TRUE)
# colnames(income_2019) = gsub('Estimate!!Number!!NONFAMILY HOUSEHOLDS!!', '', colnames(income_2019), fixed=TRUE)
# colnames(income_2019) = gsub('Estimate!!Number!!', '', colnames(income_2019), fixed=TRUE)
# head(income_2019)
# head(WI_data)
```



This dataset includes information about Public Libraries in Wisconsin. It includes information about funding, location, hours, resources and how the public utilizes the library. 
```{r, include=FALSE}
# library(readxl)
# libraries_2019 <- read_excel("PRELIMINARY_2019_public_library_service_data.xlsx", na = "-", skip = 1)[-1,]
# head(libraries_2019)
```